### 2013-07-23 branching-transit.tex
WorkingWiki importation

\item This induces a distribution on $T$, the time that a successful excursion took to cross:
assuming, as is asymptotically correct, that the number of arriving lineages is independent of how long it took to get there,
if $\tau_x$ is the random time of arrival of a successful lineage,
\begin{align} \label{eqn:arrival_time_density}
 \P\{\tau_x \in dt\} = C_{s_d,x} \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
\end{align}
where $C_{s_d,x}$ is a constant depending only on $s_d$ and $x$.

\item The length of the hitchhiking segment on \emph{one side} of the selected site $L$ is exponential with rate depending on $T$;
integrating over $T$ we get
(using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
\begin{align}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] \\
 & = C_{s_d,x} \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
 &= C_{s_d,x} \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ; \label{eqn:first_integral}
\end{align}
hence, defining $\alpha = - \log(1-s_d) \approx s_d$ (note $\alpha>0$),
\begin{align}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] = \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
\end{align}
This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential with rate $x\sqrt{2}$ conditioned to be at least $\sqrt{\alpha}$.
Note that the total length of a hitchhiking segment is the sum of two independent copies of this.

@@

Suppose we have branching processes beginning at the origin at some constant rate $\lambda$ in time.
Then the occupation density at distance $x$, i.e.\ the expected number of particles found at location $x$ at any given time, at stationarity,
is $\lambda \int_0^\infty \E[Z_t(dx)] dt$, where $Z_t$ is the branching process.
Since spatial motion is independent of branching,
\begin{align}
  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) ,
\end{align}
and so using G\&R 3.471.9 and 8.468 again,
\begin{align} \label{eqn:bp_occupation}
  \int_0^\infty \E[Z_t(dx)] dt  &= \int_0^\infty e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dt \\
  &= \begin{cases}
    \sqrt{ \frac{ 2 x }{ \pi \sigma^{3} \sqrt{2\alpha} } }  K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) \qquad & d=1 \\
    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) 
    \sim \frac{ 1 }{ \sqrt{ \sigma^3  2\pi x \sqrt{2\alpha} } } \exp\left(-\frac{x\sqrt{2\alpha}}{\sigma}\right) \qquad & d=2  ,
\end{cases} 
\end{align}
where $K_0(z)$ is the modified Bessel function of the second kind of order 0.
Note that this is the Green function for a Brownian motion killed at rate $\alpha$.

We can combine this with the formula \eqref{eqn:eqfreq} for the equilibrium frequency $q(r) \approx C \exp( -\sqrt{2 \alpha} r / \sigma)$
to estimate $\lambda$, the effective outflux from the occupied patch: equating $\rho q(r)$ with $\lambda$ multiplied by equation \eqref{eqn:bp_occupation}, we get
\begin{align} \label{eqn:outflux_from_patch}
    \lambda = \text{(outflux from old patch)} = \begin{cases}
        C \rho \sigma \sqrt{2\alpha} \quad & d=1 \\
    C \rho \sigma^{3/2} (2\alpha)^{1/4} \sqrt{2 \pi x} \quad & d=2 
    \end{cases}
\end{align}
Note that this differs from the value $C\rho$ which we might have guessed.

@@

We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
\begin{align}
    (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \label{eqn:heuristic1a} \\
    \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \label{eqn:heuristic1b} \\
    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
\end{align}
The subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
the number of (extra) offspring per subcritical individual is $s_B+s_m$;
and the probability of establishment of each such extra offspring in the patch is $2s_B/\xi^2$.
The effective outflux of branching processes is computed in equation \eqref{eqn:outflux_from_patch}.
The number of establishment events per BP is as follows, in averages: 
with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
each produce $1/s_m$ total core offspring;
each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
i.e.
\begin{align}
  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{2\alpha} x}{\sigma}\right) .
\end{align}

However, this formulation counts separately each of possible multiple establishment events by a single transiting family.
Writing
\begin{align} \label{eqn:heuristic2}
  \begin{split} (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split},
\end{align}
for $\migrate$ we actually want to omit the last term.
Using the probability of establishment of a distant branching process is in equation~\eqref{eqn:bp_prob_estab}, this is
\begin{align}
  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
           &= \begin{cases}
    C \rho \sqrt{2 \alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{2\alpha}x}{\sigma} \right)  \qquad & d=1 \\
    C \rho \sqrt{2 \alpha} \times \frac{ 2 s_b \sqrt{2 \alpha A} }{ \xi^2 \pi \sigma^3} K_1\left(\frac{x\sqrt{2\alpha}}{\sigma}\right)  \qquad & d=2 .
\end{cases}
\end{align}
Finally, combining \eqref{eqn:heuristic1b} and \eqref{eqn:heuristic2}, we can also compute that
\begin{align}
  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{ \sqrt{s_m} } .
\end{align}

>>>>

\item This induces a distribution on $\tau_x$, the time that a successful excursion took to cross:
assuming, as is asymptotically correct, that the number of arriving lineages is independent of how long it took to get there,
if $\tau_x$ is the random time of arrival of a successful lineage,
\begin{align} \label{eqn:arrival_time_density}
 \P\{\tau_x \in dt\} = C_{s_d,x} \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
\end{align}
where $C_{s_d,x}$ is a constant depending only on $s_d$ and $x$.

\item The length of the hitchhiking segment on \emph{one side} of the selected site $L$ is exponential with rate depending on $\tau_x$;
integrating over $\tau_x$ we get
(using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
\begin{align}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] \\
 & = C_{s_d,x} \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
 &= C_{s_d,x} \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ; \label{eqn:first_integral}
\end{align}
hence, defining $\alpha = - \log(1-s_d) \approx s_d$ (note $\alpha>0$),
\begin{align}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] = \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
\end{align}
This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential with rate $x\sqrt{2}$ conditioned to be at least $\sqrt{\alpha}$.
Note that the total length of a hitchhiking segment is the sum of two independent copies of this.

@@

Suppose we have branching processes beginning at the origin at some constant rate $\lambda$ in time.
Then the occupation density at distance $x$, i.e.\ the expected number of particles found at location $x$ at any given time, at stationarity,
is $\lambda \int_0^\infty \E[Z_t(dx)] dt$, where $Z_t$ is the branching process.
Since spatial motion is independent of branching,
\begin{align}
  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dx ,
\end{align}
and so using G\&R 3.471.9 and 8.468 again,
\begin{align} \label{eqn:bp_occupation}
  \int_0^\infty \E[Z_t(dx)] dt  &= \int_0^\infty e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dt \\
  &= \begin{cases}
    \sqrt{ \frac{ 2 x }{ \pi \sigma^{3} \sqrt{2\alpha} } }  K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) \qquad & d=1 \\
    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) 
    \sim \frac{ 1 }{ \sqrt{ \sigma^3  2\pi x \sqrt{2\alpha} } } \exp\left(-\frac{x\sqrt{2\alpha}}{\sigma}\right) \qquad & d=2  ,
\end{cases} 
\end{align}
where $K_0(z)$ is the modified Bessel function of the second kind of order 0.
Note that this is the Green function for a Brownian motion killed at rate $\alpha$.

We can combine this with the formula \eqref{eqn:eqfreq} for the equilibrium frequency $q(r) \approx C \exp( -\sqrt{2 \alpha} r / \sigma)$
to estimate $\lambda$, the effective outflux from the occupied patch: equating $\rho q(r)$ with $\lambda$ multiplied by equation \eqref{eqn:bp_occupation}, we get
\begin{align} \label{eqn:outflux_from_patch}
    \lambda = \text{(outflux from old patch)} = \begin{cases}
        C \rho \sigma \sqrt{2\alpha} \quad & d=1 \\
    C \rho \sigma^{3/2} (2\alpha)^{1/4} \sqrt{2 \pi x} \quad & d=2 
    \end{cases}
\end{align}
Note that this differs from the value $C\rho$ which we might have guessed.
NOT RIGHT in $d=2$: SHOULD NOT DEPEND ON $x$.

@@

We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
\begin{align}
    (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \label{eqn:heuristic1a} \\
    \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \label{eqn:heuristic1b} \\
    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
\end{align}
The subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
the number of (extra) offspring per subcritical individual is $s_p+s_m$;
and the probability of establishment of each such extra offspring in the patch is $2s_p/\xi^2$.
The effective outflux of branching processes is computed in equation \eqref{eqn:outflux_from_patch}.
The number of establishment events per BP is as follows, in averages: 
with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
each produce $1/s_m$ total core offspring;
each of these produce $s_p+s_m$ extra offspring which each have chance $2s_p/\xi^2$ of establishing,
i.e.
\begin{align}
  (\mbox{\# estab events per BP}) = \frac{(s_p+s_m) 2 s_p }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{2\alpha} x}{\sigma}\right) .
\end{align}

However, this formulation counts separately each of possible multiple establishment events by a single transiting family.
Writing
\begin{align} \label{eqn:heuristic2}
  \begin{split} (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split},
\end{align}
for $\migrate$ we actually want to omit the last term.
Using the probability of establishment of a distant branching process is in equation~\eqref{eqn:bp_prob_estab}, this is
\begin{align}
  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
           &= \begin{cases}
    C \rho \sqrt{2 \alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{2\alpha}x}{\sigma} \right)  \qquad & d=1 \\
    C \rho \sqrt{2 \alpha} \times \frac{ 2 s_b \sqrt{2 \alpha A} }{ \xi^2 \pi \sigma^3} K_1\left(\frac{x\sqrt{2\alpha}}{\sigma}\right)  \qquad & d=2 .
\end{cases}
\end{align}
Finally, combining \eqref{eqn:heuristic1b} and \eqref{eqn:heuristic2}, we can also compute that
\begin{align}
  (\mbox{\# estab events per successful BP}) = \frac{s_p+s_m}{ \sqrt{s_m} } .
\end{align}
